---
title: "terrain_gen"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Install required package

```{r}
#install.packages("data.tree")
#install.packages("tidyverse")
library("data.tree")
library("tidyverse")
```

## Naive Particle Filter

```{r}
# Read data
dataset <- read.csv("simulated_dataset.csv")
terrain_map <- dataset$terrain_map
noiseless_measurements <- dataset$noiseless_measurements[!is.na(dataset$noiseless_measurements)]
measurements <- dataset$measurements[!is.na(dataset$measurements)]

ground_truth_pos <- dataset$true_aircraft_pos[!is.na(dataset$true_aircraft_pos)]
# Read config
config <- read.csv("config.csv")
# min_v, max_v, aircraft_height, ra_noise_mean, ra_noise_sd
```

### Visualize data
```{r}
plot(terrain_map)
plot(measurements)
print(sprintf("Total measurement: %d", length(measurements)))
```

### Result evaluation metric

```{r}
# scaled MSE
evaluate <- function(predicted_pos, true_pos = ground_truth_pos) {
    # trim longer vector to the shorter one
    # Better heuristic can be used to improve evaluation metric...
    # e.g. linear regression to add missing value is not a bad idea.
    target_len <- min(c(length(true_pos), length(predicted_pos)))
    real_vec <- true_pos[1:target_len]
    obs_vec <- predicted_pos[1:target_len]
    diff_vec <- real_vec - obs_vec
    sum(diff_vec^2) / (length(diff_vec))^2 # MSE
}

mse <- function(predicted_pos, true_pos = ground_truth_pos) {
    evaluate(predicted_pos, true_pos)
}

aligned_mse <- function(predicted_pos, true_pos = ground_truth_pos) {
    start_ind <- 1
    while (start_ind < length(true_pos) && predicted_pos[1] > true_pos[start_ind]) {
        start_ind <- start_ind + 1
    }
    if (start_ind == length(true_pos)) {
        10000000 # Totally off; should not happen
    } else {
        trimmed_true_pos <- true_pos[start_ind:length(true_pos)]
        mse(predicted_pos, trimmed_true_pos)
    }
}
```

### Utility function
```{r}
argmax <- function (vec) {which.max(vec)}

calc_likelihood <- function(measured_dist, low_ind = 1, high_ind = length(terrain_map)) {
    if (low_ind > high_ind || high_ind > length(terrain_map)) {
        print("Hey")
        stop("Index error")
    }
    noise_mean <- config$ra_noise_mean
    noise_sd <- config$ra_noise_sd
    aircraft_height <- config$aircraft_height
    
    height_terrain_below <- aircraft_height - measured_dist
    # vectorized calculation to compute likelihood at each spot
    likelihood_vec <- dnorm(height_terrain_below, terrain_map[low_ind:high_ind] + noise_mean, sd = noise_sd)
    likelihood_vec
}
```

## MLE estimator

Assume that no prior knowledge was given. That is, we don't know the initial position of the aircraft
or the range of velocity that the aircraft is capable of traveling. In this case, a common option to estimate
the position would be Maximum Likelihood Estimator (MLE).

First assume that we have an ideal radio altimeter. In this case, every measurement is accurate.

### MLE with no noise

```{r}
predicted_vec <- numeric(length(measurements))
for (i in 1:length(predicted_vec)) {
    predicted_vec[i] <- argmax(calc_likelihood(noiseless_measurements[i]))
}
plot(predicted_vec)
print(sprintf("MSE with noiseless measurements: %f", evaluate(predicted_vec)))
```

Observe that we have zero Mean Squared Error (MSE). That means, we can determine the pricise coordinate
of our aircraft without any sort of prior knowledge. However, real life is not perfect. There are some
noise inherent from the sensor that we use to send out the signal. We can run the same procedure against
noisy measurements.

### MLE with noise

```{r}
predicted_vec <- numeric(length(measurements))
for (i in 1:length(predicted_vec)) {
    predicted_vec[i] <- argmax(calc_likelihood(measurements[i]))
}
plot(predicted_vec)
print(sprintf("MSE with noisy measurementsk: %f", evaluate(predicted_vec)))
```

We have a very, very large MSE and the plot produced shows a diverge pattern, which implies that the 
estimation is almost meaningless. Now we may assume that we know the range of the velocity that the
aircraft is traveling. What can we do with this prior knowledge?

## Naive Markov Chain

A Markov-chain-based strategy which performs sequential updates may potentially be a good approach.
However, there is a huge problem here: we do not know the initial position of the aircraft.
We have to be extremely careful choosing one since the output is very sensitive to this.

Here we use MLE to estimate the initial position of our aircraft.

```{r}
initial_pos <- argmax(calc_likelihood(measurements[1]))
cur_pos <- initial_pos
max_pos <- length(terrain_map)
predicted_vec <- numeric()
cur_measurement_counter <- 2
while (cur_pos <= max_pos) {
    predicted_vec <- append(predicted_vec, c(cur_pos))
    possible_vel <- config$min_v:config$max_v
    candidate_next_pos <- cur_pos + possible_vel
    candidate_next_pos <- candidate_next_pos[candidate_next_pos <= length(terrain_map)]
    if (length(candidate_next_pos) == 0 || cur_measurement_counter > length(measurements)) {
        cur_pos <- max_pos + 1 # halt
    } else {
        # got some valid candidates
        candidate_len <- length(candidate_next_pos)
        candidate_likelihood <- calc_likelihood(measurements[cur_measurement_counter],
                                                low_ind = candidate_next_pos[1],
                                                high_ind = candidate_next_pos[candidate_len])
        cur_measurement_counter <- cur_measurement_counter + 1
        cur_pos <- candidate_next_pos[argmax(candidate_likelihood)]
    }
}
plot(predicted_vec)
print(sprintf("Trimmed MSE: %f", aligned_mse(predicted_vec)))
```

The output makes a lot more sense and MSE is reduced by a few order of magnitudes. Note that in our
implementation, the update to each state depends only on current state. Therefore it's a Markov chain.

Note that for simplicity, the initial position of the aircraft in the dataset was set to $1$, though we
assumed it was unknown. Also observe that from the plot we know that the MLE estimator estimates that
the aircraft starts at a location that differs very much from the true initial position. This introduced
most of the errors.

Just for fun, what if we assume that we have an oracle that tells us out initial location?
```{r}
initial_pos <- 1
cur_pos <- initial_pos
max_pos <- length(terrain_map)
predicted_vec <- numeric()
my_likelihood_vec <- c()
cur_measurement_counter <- 2
while (cur_pos <= max_pos) {
    predicted_vec <- append(predicted_vec, c(cur_pos))
    possible_vel <- config$min_v:config$max_v
    candidate_next_pos <- cur_pos + possible_vel
    candidate_next_pos <- candidate_next_pos[candidate_next_pos <= length(terrain_map)]
    if (length(candidate_next_pos) == 0 || cur_measurement_counter > length(measurements)) {
        cur_pos <- max_pos + 1 # halt
    } else {
        # got some valid candidates
        candidate_len <- length(candidate_next_pos)
        candidate_likelihood <- calc_likelihood(measurements[cur_measurement_counter],
                                                low_ind = candidate_next_pos[1],
                                                high_ind = candidate_next_pos[candidate_len])
        cur_measurement_counter <- cur_measurement_counter + 1
        my_likelihood_vec <- append(my_likelihood_vec, c(max(candidate_likelihood)))
        cur_pos <- candidate_next_pos[argmax(candidate_likelihood)]
    }
}
plot(predicted_vec)
print(sprintf("MSE: %f", mse(predicted_vec)))
```

## Particle filter with backtracking

```{r}
# Hyperparameters
resampling_size <- 300
```

```{r}
predicted_vec <- read.csv("particle_filter.csv")$x
plot(predicted_vec, type = "l", col = "red")
lines(ground_truth_pos)
mse(predicted_vec)
```